---
title: "Data viz"
author: "Beata"
date: "17/06/2021"
output: html_document
---
```{r setup}
#load packages
source(here::here("R/package-loading.R"))

#load the dataset- tided
load(here::here("data/nhanes_small.rda"))

```

#one variable plot- continous variables- nr, 
```{r continous-variables}
ggplot(nhanes_small, aes(x = bmi)) +
    geom_density()

ggplot(nhanes_small, aes(x = bmi)) +
    geom_histogram()
```
```{r discrete_variables}
ggplot(nhanes_small, aes(x = sex)) +
    geom_bar()

# add diabets status
ggplot(nhanes_small, aes(x = diabetes)) +
    geom_bar()
```
#remove NA from graph- data wrangling- new data set called nhanes_tidied done fron nhanes_small
```{r remove missing-diabetes}
nhanes_tidied <- nhanes_small %>% 
    filter(!is.na(diabetes))

#safe new dataset that can be edited-overwrite again in the code above ...add stuff to code- remove NA fot other value- do not need to safe it with new name. 
usethis::use_data(nhanes_tidied, overwrite = T)

#new graph with NA removed- nhanes tided
ggplot(nhanes_tidied, aes(x = diabetes)) +
    geom_bar()
```

## PLOTTING 2 VARIABLES with new file name bmi_chol <- (this is to make many diffrent plots from that without writing ---ggplot(nhanes_tidied, aes (x = bmi, y = tot_chol)) EVERY TIME), assign the as object again - run it first before running the plot!- new file in the Environment 

```{r 2-continious-variables}
bmi_chol <- ggplot(nhanes_tidied, aes (x = bmi, y = tot_chol))

#so now we can just write bmi_chol + ... geom_... (because the bmi_chol + is safed <-... the long function)

bmi_chol +
    geom_point()

bmi_chol +
    geom_hex()

# doubble graph

bmi_chol +
geom_hex () +
geom_smooth()
```

# assign the as object again - run it first before running the plot!
```{r 2-disrete-variables}
two_discrete <- nhanes_tidied %>% 
    ggplot(aes(x = diabetes, fill = sex))

two_discrete +
    geom_bar()

# not on top of each other but side by side

two_discrete +
    geom_bar(position = position_dodge())
```
```{r discrete-continous-variables}
#create and object <-  

tow_mixed <- nhanes_tidied %>% 
    ggplot(aes(x = diabetes, y = bmi))

#then make a graph

tow_mixed +
    geom_boxplot()

# spotts graph

tow_mixed +
    geom_jitter()

#ray plot

tow_mixed +
    geom_violin()

```
